Perceived value of video games, but not hours played, predicts mental well-being in adult Nintendo players

### 2024-06-26 ms.qmd
fix table and fig 2

library(tidyverse)
library(lubridate)
library(knitr)
library(kableExtra)
library(dtplyr)
library(scales)
library(ggdist)
library(distributional)
library(mgcv)
library(patchwork)
library(see)
library(performance)
library(careless)
library(cowplot)
library(marginaleffects)
library(flextable)

@@

ethnicities <- datDemo |> 
  filter(pid %in% dat$pid & ethnicity != "DATA_EXPIRED") |> 
  count(ethnicity, sort = TRUE) |>
  mutate(p = n / sum(n)) |>
  mutate(x = str_glue("{ethnicity} ({number(n, big.mark = '')}, {percent(p)})")) |>
  slice(1:4) |>
  pull(x) |>
  knitr::combine_words()

jobs <- datDemo |> 
  filter(pid %in% dat$pid & employmentStatusProlific != "DATA_EXPIRED") |> 
  count(employmentStatusProlific, sort = TRUE) |>
  mutate(p = n / sum(n)) |>
  mutate(x = str_glue("{employmentStatusProlific} ({number(n, big.mark = '')}, {percent(p)})")) |>
  slice(1:4) |>
  pull(x) |>
  knitr::combine_words()

dat |> 
  select(age, gender) |> 
  mutate(
    Gender = ifelse(!(gender %in% c("Man","Woman")), 
                     "Non-binary or other gender identity", 
                     gender)
  ) |> 
  group_by(Gender) |> 
  summarise(
    n = n(),
            `Median age (SD)` = paste0(median(age, na.rm = TRUE), " (", round(sd(age, na.rm = TRUE), 2), ")")
    ) |> 
  flextable(cwidth = c(2.5, 1, 1)) %>% 
  theme_apa()

Demographic information about participants is shown in @tbl-demographics. All participants were residents of the US, and were majority white: `r ethnicities`. Approximately half were in full-time employment: `r jobs`.

@@

```{r}
#| label: fig-playtime
#| fig-cap: "Density of playtime across each of the observation periods from 1 hour to 1 year."
#| fig-width: 8
#| fig-height: 6
#| fig-align: center
#| fig-pos: "t"
#| apa-twocolumn: true

# played <- dat |> 
#   select(pid, matches("played\\d")) |> 
#   pivot_longer(cols = -pid, names_to = "timePeriod", values_to = "played") %>%
#   group_by(timePeriod) %>%
#   summarise(
#     count = sum(played),
#     prop = sum(played, na.rm = TRUE)/ n(),
#   ) |> 
#   mutate(
#     timePeriod = factor(timePeriod, levels = playedDict),
#     timePeriod = fct_rev(fct_recode(timePeriod, !!!playedDict)),
#   ) |> 
#   arrange(timePeriod) |> 
#   mutate(index = 1:n()) |> 
#   ggplot(aes(y = timePeriod, x = prop)) +
#   geom_bar(stat = "identity", fill = "#164187") + 
#   geom_text(aes(label = paste0(round(prop*100, 0), "% (", count, ")"),
#                 hjust = ifelse(index <= 4, 1.1, -.1),
#                 color = index <= 4), 
#             size = 4) +
#   mytheme + 
#   theme(plot.margin = margin(3,1.5,3,0, "cm"),
#         legend.position = "none") +
#   scale_x_continuous(breaks = seq(0, 1, .25), labels = scales::percent) +
#   scale_color_manual(values = c("black","white")) +
#   labs(fill = "", x = "Proportion of\nparticipants with\nany playtime", y = "")
# 
# playtimeHist <- dat |> 
#   filter(!is.na(recordedDate)) |> 
#   select(pid, recordedDate, matches("playtime\\d")) |> 
#   pivot_longer(cols = starts_with("playtime"), names_to = "timePeriod") |>
#   filter(
#     value > 0 & # we plot average playtime only for players with some logged data
#       value < 2 # for some reason, we need to set the xmax limit here, as filtering later causes issues with proportion
#     ) |> 
#   mutate(
#     timePeriod = factor(timePeriod, levels = playtimeDict),
#     timePeriod = fct_recode(timePeriod, !!!playtimeDict)
#   ) |> 
#   mutate(value = value*60) |> # convert to minutes for a more granular plot
#   ggplot(
#     aes(x = value, 
#         fill = timePeriod, 
#         group = timePeriod,
#         y = after_stat(density*width)) # width * density gives proper proportions in facet_wrapped plots
#   ) + 
#   geom_histogram(binwidth = .2*60, color = "black") +
#   scale_y_continuous(limits = c(0, .65), labels = scales::percent) +
#   scale_x_continuous(limits = c(-10, 120), breaks = seq(0, 120, 30)) +
#   scale_fill_viridis_d() +
#   mytheme + 
#   labs(
#     x = "Minutes of Daily Play\n(where playtime window < 24 hours,\nraw minutes)",
#     y = "Proportion of responses"
#   ) + 
#   theme(legend.position = "none",
#         axis.text.x = element_text(size = 10, angle = 30, vjust = 1, hjust = 1)) +
#   facet_wrap(~timePeriod)
# 
# plot_grid(played, playtimeHist, ncol = 2, rel_widths = c(4, 6))

#### Heatmap attempt
dat |> 
  select(pid, recordedDate, matches("playtime\\d")) |> 
  pivot_longer(cols = starts_with("playtime"), names_to = "timePeriod") |>
  filter(!is.na(value)) |> 
  mutate(
    timePeriod = factor(timePeriod, levels = playtimeDict),
    timePeriod = fct_recode(timePeriod, !!!playtimeDict)
  ) |> 
  mutate(
    playtimeCat = factor(case_when(
    value > 40/60 ~ ">40\nmin/day",
    value > 30/60 ~ "30-40\nmin/day",
    value > 20/60 ~ "20-30\nmin/day",
    value > 10/60 ~ "10-20\nmin/day",
    value > 0 ~ "0-10\nmin/day",
    value == 0 ~ "No play during\nobservation\nperiod"),
    levels = c("No play during\nobservation\nperiod", "0-10\nmin/day", "10-20\nmin/day", "20-30\nmin/day", 
               "30-40\nmin/day", ">40\nmin/day"))
  ) |> 
  group_by(playtimeCat, timePeriod) |> 
  summarise(count = n(),
            prop = n()/nrow(dat)) |> 
  ungroup() |> 
  complete(timePeriod, playtimeCat, fill = list(count = 0, prop = 0)) |> 
  ggplot(aes(playtimeCat, timePeriod, fill = count)) + 
  geom_tile() +
  geom_text(aes(label = count), color = "#404040") +
  labs(x = "", y = "Observation Window", fill = "Count") +
  scale_fill_gradient(low="white", high="blue") +
  theme_half_open()

dat |> 
  select(playtime1d, playtime2wk, playtime3mo, playtime1yr) |> 
  # convert back to total playtime in hours, rather than minutes per day
  # mutate(
  #   playtime1d = playtime1d*24,
  #   playtime2wk = playtime2wk*24*14,
  #   playtime3mo = playtime3mo*24*91,
  #   playtime1yr = playtime1yr*24*36
  # ) |> 
  # scale_x_continuous(labels = scales::cut_si()) +
  pivot_longer(cols = everything(), 
               names_to = "timewindow", 
               names_pattern = "playtime(.*)", 
               values_to = "playtime") |> 
  mutate(timewindow = fct_recode(timewindow,
                                 "1 Day" = "1d",
                                 "2 Weeks" = "2wk",
                                 "3 Months" = "3mo",
                                 "1 Year" = "1yr"),
         timewindow = fct_relevel(timewindow, 
                                  "1 Day", 
                                  "2 Weeks", 
                                  "3 Months", 
                                  "1 Year"),
         # playtime = if_else(playtime == 0, map_dbl(playtime, ~ sample(c(0, -.1, -.2, -.3), 1)), playtime)
  ) |> 
  ggplot(aes(x = timewindow, y = playtime, fill = timewindow)) +
  stat_slab(
    density = ggdist::density_bounded(bandwidth = "nrd0"),
    normalize = "all",
    # height = 2
    scale = .4,
  ) +
  stat_dots(
    side = "left",
    justification = 1.2,
    overflow = "compress",
    # overflow = "multiple",
    binwidth = .1,
    dotsize = .5,
    scale = .4,
    color = NA,
    layout = "hex",
    # stackratio = .5,
    overlaps = "nudge",
    # verbose = TRUE,
    alpha = .3
  ) +
  geom_boxplot(
    width = .10, 
    outlier.shape = NA
  ) +
  labs(x = "Observation Period", y = "Daily Hours of Playtime") +
  scale_y_continuous(
    trans = "sqrt",
    breaks = c(0, .2, .5, 1, 2, 3, 4, 5),
    limits = c(0, 5)
    # limits = c(-.3, 5),
  ) +
  scale_x_discrete(expand = c(0, 0)) +
  coord_cartesian(xlim = c(.3, 4.3)) +
  mytheme +
  theme(legend.position = "none") +
  scale_fill_viridis_d()

>>>>

library(tidyverse)
library(lubridate)
library(knitr)
library(kableExtra)
library(dtplyr)
library(scales)
library(ggdist)
library(distributional)
library(mgcv)
library(patchwork)
library(see)
library(performance)
library(careless)
library(cowplot)
library(marginaleffects)
library(flextable)
library(ggstance)
library(crosstable)

@@

dat |> 
  left_join(datDemo, by = "pid") |> 
  select(age, gender, employment, eduLevel, ethnicity) |> 
  mutate(
    Age = age,
    Gender = ifelse(!(gender %in% c("Man","Woman")), 
                     "Non-binary or other gender identity", 
                     gender),
    `Employment Status` = ifelse(!(employment %in% c("Working full-time","Working part-time", "Student", "Not currently employed")), 
                     "Other employment status", 
                     employment),
    Ethnicity = ifelse(!(ethnicity %in% c("White","Asian","Mixed","Black")), 
                     "Other ethnicity", 
                     ethnicity),
    `Educational Level` = ifelse(!(eduLevel %in% c("University Bachelors Degree","Asian","Mixed","Black")), 
                     "Other ethnicity", 
                     eduLevel)
  ) |> 
  mutate(
    Gender = fct_relevel(Gender, sort(table(Gender), decreasing = TRUE) %>% names()),
    `Employment Status` = fct_relevel(`Employment Status`, sort(table(`Employment Status`), decreasing = TRUE) %>% names()),
    Ethnicity = fct_relevel(Ethnicity, sort(table(Ethnicity), decreasing = TRUE) %>% names()),
    `Educational Level` = fct_relevel(`Educational Level`, sort(table(`Educational Level`), decreasing = TRUE) %>% names())
  ) |> 
  crosstable(
    cols = c("Age","Gender","Employment Status","Ethnicity"),
    percent_digits = 1,
    funs = c("Mean (SD)" = meansd, "Median (IQR)" = mediqr, "Min / Max" = minmax)
  ) |> 
  crosstable::as_flextable() |> 
  set_header_labels(
    values = c(
      "Descriptor", "Variable", "Value")
  ) |>
  theme_apa() |> 
  fontsize(size = 8, part = "all") |>  # Adjust font size
  padding(padding = 0, part = "all")  # Adjust padding
```

Participant demographics are shown in @tbl-demographics.

@@

```{r}
#| label: fig-playtime
#| fig-cap: "Description of playtime in the sample. Panel A shows the distribution of playtime across players for each observation period, with the mean shown in red; observations periods below the dashed line are shown as total hours, whereas periods above the dashed line reflect hours of player per day. Panel B shows the proportion of players who logged at least one play session in that period."
#| fig-width: 8
#| fig-height: 6
#| fig-align: center
#| fig-pos: "t"
#| apa-twocolumn: true

p1 <- dat |> 
  select(pid, matches("playtime\\d")) |> 
  pivot_longer(-pid) |> 
  mutate(name = fct_rev(fct_recode(name, !!!playtimeDict))) |> 
  mutate(name = fct_inorder(name)) |> 
  ggplot(aes(value, name)) +
  geom_point(alpha = .6) +
  geom_hline(yintercept = 5.5, linetype = "dashed") +
  stat_summary(fun = mean, col = "red") +
  labs(
    x = "Hours of (daily) play",
    y = "Observation Period"
  ) +
  scale_x_continuous(limits = c(0, 6), 
                     breaks = c(0, 1, 2, 3, 4, 5, 6)) +
  mytheme

p2 <- dat |> 
  select(pid, matches("played\\d")) |> 
  mutate(across(where(is.logical), as.numeric)) |> 
  pivot_longer(-pid) |> 
  mutate(name = fct_rev(fct_recode(name, !!!playedDict))) |> 
  mutate(name = fct_inorder(name)) |> 
  summarise(
    n = sum(value),
    .by = c(name)
  ) |> 
  ggplot(aes(n, name)) +
  scale_x_continuous(
    "Participants",
    sec.axis = sec_axis(
      ~ . / length(unique(dat$pid)), name = "Proportion"
    )
  ) +
  labs(y = "") +
  geom_point() +
  mytheme

(p1 | p2) + plot_layout(axes = "collect") + plot_annotation(tag_level = "A")
